Analyzing Performance Bottleneck in Graph Neural Network Training: An Experimental View

### 2020-07-14 Manuscript/sec2-review-of-graph-neural-network.tex
更新Section2

\begin{figure}
	\centering
	\includegraphics[width=1.1\columnwidth]{figs/illustration/GNN_common_architecture.png}
	\caption{Structure of a typical graph neural network. (a) Demo GNN, (b) Demo graph. The target application is the node classification. The demo GNN has two GNN layers.}
	\label{fig:general_structure_of_gnn}
\end{figure}

@@

\begin{table}
	\hspace{-10em}
    \begin{footnotesize}
        \begin{tabular}{ccp{8em}p{22em}r}
		\toprule
		GNN &
		Type &
		$\Sigma$ &
		$\phi$ &
		Complexity \\ \midrule
		ChebNet \cite{defferrad2016_chebnet} &
		Spectral &
		sum &
		$\boldsymbol{m}_{j, i}^{(k)} = e_{j, i}\boldsymbol{z}_j^{(k-1)}$ &
		$O(d_{in})$ \\
		\textbf{GCN} \cite{kipf2017_gcn} &
		Spectral &
		sum &
		$\boldsymbol{m}_{j, i}^l = e_{j, i} \boldsymbol{h}_j^l$ &
		$O(d_{in})$ \\
		AGCN &
		Spectral &
		sum &
		$\boldsymbol{m}_{j, i}^l = \tilde{e}_{j, i}^l \boldsymbol{h}_j^l$ &
		$O(d_{in})$ \\
		GraphSAGE &
		Non-spectral &
		mean/LSTM &
		$\boldsymbol{m}_{j, i}^l =  \boldsymbol{h}_j^l$ &
		$O(1)$ \\
		GraphSAGE-pool &
		Non-spectral &
		max &
		$\boldsymbol{m}_{j, i}^l =  \delta(\textcolor{blue}{\boldsymbol{W}^l_{pool}} \boldsymbol{h}_j^l + \textcolor{blue}{b}^l)$  &
		$O(d_{in} * d_{out})$ \\
		Neural FPs &
		Non-spectral &
		sum &
		$\boldsymbol{m}_{j, i}^l = \boldsymbol{h}_j^l$ &
		$O(1)$ \\
		SSE &
		Recurrent &
		sum &
		$\boldsymbol{m}_{j, i}^l = [\boldsymbol{h}_i^{l} \parallel \boldsymbol{h}_j^l]$ &
		$O(d_{in})$ \\
		\textbf{GGNN} &
		Gated &
		sum &
		$\boldsymbol{m}_{j, i} = \textcolor{blue}{\boldsymbol{W}^l} \boldsymbol{h}_j^l$ &
		$O(d_{in} * d_{out})$ \\
		\textbf{GAT}&
		Attention &
		sum  &
		\begin{scriptsize}
			$\begin{aligned}[t]
			& \alpha_{j, i}^k = \frac {\exp(LeakyReLU(\textcolor{blue}{\boldsymbol{a}^T} [ \textcolor{blue}{\boldsymbol{W}^{l,k}}  \boldsymbol{h}_j^l \parallel \textcolor{blue}{\boldsymbol{W}^{l,k}}  \boldsymbol{h}_j^l] ))} {\sum_{k \in \mathcal{N}(i)}\exp(LeakyReLU(\textcolor{blue}{\boldsymbol{a}^T} [ \textcolor{blue}{\boldsymbol{W}^{l,k}}  \boldsymbol{h}_j^l \parallel \textcolor{blue}{\boldsymbol{W}^{l,k}}  \boldsymbol{h}_k^l] ))} \\
			& \text{Multi-head concatenation}: \boldsymbol{m}_{j, i}^l = \parallel_{k=1}^K \delta(\alpha_{j, i}^k \textcolor{blue}{\boldsymbol{W}^{l,k}} \boldsymbol{h}_j^{l}) \\
			& \text{Multi-head average}: \boldsymbol{m}_{j, i}^l = \frac{1}{K} \sum_{k=1}^K \delta(\alpha_{j, i}^k \textcolor{blue}{\boldsymbol{W}^{l,k}} \boldsymbol{h}_j^{l})
			\end{aligned}$
		\end{scriptsize} &
		$O(K * d_{in} * d_{out})$ \\
		\textbf{GaAN} &
		Attention &
		sum,max,mean &
		\begin{scriptsize}
			$\begin{aligned}[t]
			&\alpha_{j, i}^k = \frac {\exp(\textcolor{blue}{\boldsymbol{a}^T} [ \textcolor{blue}{\boldsymbol{W}^{l, k}_{xa}} \boldsymbol{h}_j^l \parallel \textcolor{blue}{\boldsymbol{W}^{l,k}_{ya}} \boldsymbol{h}_i^l] )} {\sum_{k \in \mathcal{N}(j)}\exp(\textcolor{blue}{\boldsymbol{a}^T} [ \textcolor{blue}{\boldsymbol{W}^{l,k}_{xa}} \boldsymbol{h}_j^l \parallel \textcolor{blue}{\boldsymbol{W}^{l,k}_{ya}}  \boldsymbol{h}_k^l] )} \\
			&\boldsymbol{m}_{j, i, 1}^l = \parallel_{k=1}^K \delta(\alpha_{j, i}^k \textcolor{blue}{\boldsymbol{W}^{l,k}_v} \boldsymbol{h}_j^{l}) \\
			&\boldsymbol{m}_{j, i, 2}^l = \textcolor{blue}{\boldsymbol{W}^l_m} \boldsymbol{h}_j^{l} \\
			&\boldsymbol{m}_{j, i, 3}^l = \boldsymbol{h}_j^l
			\end{aligned}$
		\end{scriptsize} &
		$O(max(d_a, d_m, d_v) * K * d_{in})$ \\
		\bottomrule
	\end{tabular}
\end{footnotesize}
	\caption{Typical graph neural networks and their edge computation functions.$d_{in}$ and $d_{out}$ are dimensions of the input and output hidden feature vectors, respectively.
		Blue variables are model parameters to learn. For ChebNet, we report its GNN sub-layer. Because two sum operator are interchangeable and sum operator and concatenation operator are interchangeable, we do it for GAT to explain more clearly in edge computations functions.}
	\label{tab:gnn_overview_edge}
\end{table}

\begin{table}
	\hspace{-3em}
    \begin{footnotesize}
        \begin{tabular}{cp{20em}r}
		\toprule
		GNN &
		$\gamma$ &
		Complexity \\ \midrule
		ChebNet \cite{defferrad2016_chebnet} &
		$\boldsymbol{z}_i^{(k)} = 2\boldsymbol{s}^{(k)}_{i} - \boldsymbol{z}_i^{(k-2)}$&
		$O(d_{in})$ \\
		\textbf{GCN} \cite{kipf2017_gcn} &
		$\boldsymbol{h}_i^{l+1} = \textcolor{blue}{\boldsymbol{W}}^l  \boldsymbol{s}_i^{l}$ &
		$O(d_{in} * d_{out})$ \\
		AGCN &
		$\boldsymbol{h}_i^{l+1} = \textcolor{blue}{\boldsymbol{W}}^l  \boldsymbol{s}_i^{l}$ &
		$O(d_{in} * d_{out})$ \\
		GraphSAGE &	
		$\boldsymbol{h}_i^{l+1} =   \delta(\textcolor{blue}{\boldsymbol{W}}^l  [\boldsymbol{s}_i^{l} \parallel \boldsymbol{h}_i^l])$ &
		$O(d_{in} * d_{out})$ \\
		GraphSAGE-pool &	
		$\boldsymbol{h}_i^{l+1} = \boldsymbol{s}_i^l$ &
		$O(1)$ \\
		Neural FPs &
		$\boldsymbol{h}_i^{l+1} = \delta(\textcolor{blue}{\boldsymbol{W}}^{l, |\mathcal{N}(i)|}  (\boldsymbol{h}_i^l + \boldsymbol{s}_i^{l}))$ &
		$O(d_{in} * d_{out})$ \\
		SSE &
		$\boldsymbol{h}_i^{l+1} = (1 - \alpha)  \boldsymbol{h}_i^l +\alpha    \delta(\textcolor{blue}{\boldsymbol{W}}^l_1 \delta(\textcolor{blue}{\boldsymbol{W}}^l_2 \boldsymbol{s}_i^l))$ &
		$O(d_{in} * d_{out})$ \\
		\textbf{GGNN} &
		\begin{scriptsize}
		$\begin{aligned}[t]
			& \textcolor{blue}{\boldsymbol{z}}_i^l = \delta ( \textcolor{blue}{\boldsymbol{W}}^z \boldsymbol{s}_i^l + \textcolor{blue}{\boldsymbol{b}}^{sz} + \textcolor{blue}{\boldsymbol{U}}^z \boldsymbol{h}_i^{l} + \textcolor{blue}{\boldsymbol{b}}^{hz}) \\ 
			&\boldsymbol{r}_i^l = \delta ( \textcolor{blue}{\boldsymbol{W}}^r \boldsymbol{s}_i^l+ \textcolor{blue}{\boldsymbol{b}}^{sr} +\textcolor{blue}{\boldsymbol{U}}^r \boldsymbol{h}_i^{l} + \textcolor{blue}{\boldsymbol{b}}^{hr}) \\ 
			&\boldsymbol{h}_i^{l+1} = tanh ( \textcolor{blue}{\boldsymbol{W}} \boldsymbol{s}_i^l + \textcolor{blue}{\boldsymbol{b}}^s + \textcolor{blue}{\boldsymbol{U}} ( \boldsymbol{r}_i^l \odot \boldsymbol{h}_i^{l}) + \textcolor{blue}{\boldsymbol{b}}^h)) \\ 
			&\boldsymbol{h}_i^{l+1} = (1 - \textcolor{blue}{\boldsymbol{z}}_i^l) \odot \boldsymbol{h}_i^l +  \textcolor{blue}{\boldsymbol{z}}_i^l \odot \boldsymbol{h}_i^{l+1}
		\end{aligned}$
		\end{scriptsize}
		&
		$O(max(d_{in}, d_{out}) * d_{out})$ \\
		\textbf{GAT}&
		$\boldsymbol{h}_i^{l+1} = \boldsymbol{s}_i^l$ &
		$O(1)$ \\
		\textbf{GaAN} &
		\begin{scriptsize}
			$\begin{aligned}[t]
			&\boldsymbol{g}_i = \textcolor{blue}{\boldsymbol{W}}^l_g  [\boldsymbol{h}_i^{l} \parallel \boldsymbol{s}_{i, 2}^l \parallel \boldsymbol{s}_{i, 3}^l]  \\
			&\boldsymbol{h}_i^{l+1} = \textcolor{blue}{\boldsymbol{W}}^l_o [\boldsymbol{h}_i^l \parallel (\boldsymbol{g}_{i} \odot \boldsymbol{s}_{i, 1}^l) ]
			\end{aligned}$ 
		\end{scriptsize}&
		$O(max(K * d_v + d_{in}, 2 * d_{in} + d_m) * d_{out})$ \\ \bottomrule
	\end{tabular}
\end{footnotesize}
	\caption{Typical graph neural networks and their vertex computation functions. $d_{in}$ and $d_{out}$ are dimensions of the input and output hidden feature vectors, respectively. Blue variables are model parameters to learn. For ChebNet, we report its GNN sub-layer.For Neural FPs, $\textcolor{blue}{\boldsymbol{W}}^{l, |\mathcal{N}(i)|}$ is the weight matrix for nodes with degree
	$|\mathcal{N}(i)|$ at layer $l$.}
	\label{tab:gnn_overview_vertex}
\end{table}

@@

>>>>

\begin{figure}
	\centering
	\includegraphics[width=1\columnwidth]{figs/illustration/GNN_common_architecture.png}
	\caption{Structure of a typical graph neural network. (a) Demo GNN, (b) Demo graph. The target application is the node classification. The demo GNN has two GNN layers.}
	\label{fig:general_structure_of_gnn}
\end{figure}

@@

\begin{table}
	\hspace{-5em}
    \begin{footnotesize}
        \begin{tabular}{ccp{8em}p{22em}r}
		\toprule
		GNN &
		Type &
		$\Sigma$ &
		$\phi$ &
		Complexity \\ \midrule
		ChebNet \cite{defferrad2016_chebnet} &
		Spectral &
		sum &
		$\boldsymbol{m}_{j, i}^{(k)} = e_{j, i}\boldsymbol{z}_j^{(k-1)}$ &
		$O(d_{in})$ \\
		\textbf{GCN} \cite{kipf2017_gcn} &
		Spectral &
		sum &
		$\boldsymbol{m}_{j, i}^l = e_{j, i} \boldsymbol{h}_j^l$ &
		$O(d_{in})$ \\
		AGCN &
		Spectral &
		sum &
		$\boldsymbol{m}_{j, i}^l = \tilde{e}_{j, i}^l \boldsymbol{h}_j^l$ &
		$O(d_{in})$ \\
		GraphSAGE &
		Non-spectral &
		mean/LSTM &
		$\boldsymbol{m}_{j, i}^l =  \boldsymbol{h}_j^l$ &
		$O(d_{in})$ \\
		GraphSAGE-pool &
		Non-spectral &
		max &
		$\boldsymbol{m}_{j, i}^l =  \delta(\textcolor{blue}{\boldsymbol{W}^l_{pool}} \boldsymbol{h}_j^l + \textcolor{blue}{b}^l)$  &
		$O(d_{in} * d_{out})$ \\
		Neural FPs &
		Non-spectral &
		sum &
		$\boldsymbol{m}_{j, i}^l = \boldsymbol{h}_j^l$ &
		$O(d_{in})$ \\
		SSE &
		Recurrent &
		sum &
		$\boldsymbol{m}_{j, i}^l = [\boldsymbol{h}_i^{l} \parallel \boldsymbol{h}_j^l]$ &
		$O(d_{in})$ \\
		\textbf{GGNN} &
		Gated &
		sum &
		$\boldsymbol{m}_{j, i} = \textcolor{blue}{\boldsymbol{W}^l} \boldsymbol{h}_j^l$ &
		$O(d_{in} * d_{out})$ \\
		\textbf{GAT}&
		Attention &
		sum  &
		\begin{scriptsize}
			$\begin{aligned}[t]
			& \alpha_{j, i}^k = \frac {\exp(LeakyReLU(\textcolor{blue}{\boldsymbol{a}^T} [ \textcolor{blue}{\boldsymbol{W}^{l,k}}  \boldsymbol{h}_j^l \parallel \textcolor{blue}{\boldsymbol{W}^{l,k}}  \boldsymbol{h}_j^l] ))} {\sum_{k \in \mathcal{N}(i)}\exp(LeakyReLU(\textcolor{blue}{\boldsymbol{a}^T} [ \textcolor{blue}{\boldsymbol{W}^{l,k}}  \boldsymbol{h}_j^l \parallel \textcolor{blue}{\boldsymbol{W}^{l,k}}  \boldsymbol{h}_k^l] ))} \\
			& \text{Multi-head concatenation}: \boldsymbol{m}_{j, i}^l = \parallel_{k=1}^K \delta(\alpha_{j, i}^k \textcolor{blue}{\boldsymbol{W}^{l,k}} \boldsymbol{h}_j^{l}) \\
			& \text{Multi-head average}: \boldsymbol{m}_{j, i}^l = \frac{1}{K} \sum_{k=1}^K \delta(\alpha_{j, i}^k \textcolor{blue}{\boldsymbol{W}^{l,k}} \boldsymbol{h}_j^{l})
			\end{aligned}$
		\end{scriptsize} &
		$O(K * d_{in} * d_{out})$ \\
		\textbf{GaAN} &
		Attention &
		sum,max,mean &
		\begin{scriptsize}
			$\begin{aligned}[t]
			&\alpha_{j, i}^k = \frac {\exp(\textcolor{blue}{\boldsymbol{a}^T} [ \textcolor{blue}{\boldsymbol{W}^{l, k}_{xa}} \boldsymbol{h}_j^l \parallel \textcolor{blue}{\boldsymbol{W}^{l,k}_{ya}} \boldsymbol{h}_i^l] )} {\sum_{k \in \mathcal{N}(j)}\exp(\textcolor{blue}{\boldsymbol{a}^T} [ \textcolor{blue}{\boldsymbol{W}^{l,k}_{xa}} \boldsymbol{h}_j^l \parallel \textcolor{blue}{\boldsymbol{W}^{l,k}_{ya}}  \boldsymbol{h}_k^l] )} \\
			&\boldsymbol{m}_{j, i, 1}^l = \parallel_{k=1}^K \delta(\alpha_{j, i}^k \textcolor{blue}{\boldsymbol{W}^{l,k}_v} \boldsymbol{h}_j^{l}) \\
			&\boldsymbol{m}_{j, i, 2}^l = \textcolor{blue}{\boldsymbol{W}^l_m} \boldsymbol{h}_j^{l} \\
			&\boldsymbol{m}_{j, i, 3}^l = \boldsymbol{h}_j^l
			\end{aligned}$
		\end{scriptsize} &
		$O(max(d_a, d_m, d_v) * K * d_{in})$ \\
		\bottomrule
	\end{tabular}
\end{footnotesize}
	\caption{Typical graph neural networks and their edge computation functions.$d_{in}$ and $d_{out}$ are dimensions of the input and output hidden feature vectors, respectively.
		Blue variables are model parameters to learn. For ChebNet, we report its GNN sub-layer. Because two sum operator are interchangeable and sum operator and concatenation operator are interchangeable, we do it for GAT to explain more clearly in edge computations functions.}
	\label{tab:gnn_overview_edge}
\end{table}

\begin{table}
	\hspace{-3em}
    \begin{footnotesize}
        \begin{tabular}{cp{20em}r}
		\toprule
		GNN &
		$\gamma$ &
		Complexity \\ \midrule
		ChebNet \cite{defferrad2016_chebnet} &
		$\boldsymbol{z}_i^{(k)} = 2\boldsymbol{s}^{(k)}_{i} - \boldsymbol{z}_i^{(k-2)}$&
		$O(d_{out})$ \\
		\textbf{GCN} \cite{kipf2017_gcn} &
		$\boldsymbol{h}_i^{l+1} = \textcolor{blue}{\boldsymbol{W}}^l  \boldsymbol{s}_i^{l}$ &
		$O(d_{in} * d_{out})$ \\
		AGCN &
		$\boldsymbol{h}_i^{l+1} = \textcolor{blue}{\boldsymbol{W}}^l  \boldsymbol{s}_i^{l}$ &
		$O(d_{in} * d_{out})$ \\
		GraphSAGE &	
		$\boldsymbol{h}_i^{l+1} =   \delta(\textcolor{blue}{\boldsymbol{W}}^l  [\boldsymbol{s}_i^{l} \parallel \boldsymbol{h}_i^l])$ &
		$O(d_{in} * d_{out})$ \\
		GraphSAGE-pool &	
		$\boldsymbol{h}_i^{l+1} = \boldsymbol{s}_i^l$ &
		$O(d_{out})$ \\
		Neural FPs &
		$\boldsymbol{h}_i^{l+1} = \delta(\boldsymbol{h}_i^l + \textcolor{blue}{\boldsymbol{W}}^{l, |\mathcal{N}(i)|}  \boldsymbol{s}_i^{l})$ &
		$O(d_{in} * d_{out})$ \\
		SSE &
		$\boldsymbol{h}_i^{l+1} = (1 - \alpha)  \boldsymbol{h}_i^l +\alpha    \delta(\textcolor{blue}{\boldsymbol{W}}^l_1 \delta(\textcolor{blue}{\boldsymbol{W}}^l_2 \boldsymbol{s}_i^l))$ &
		$O(d_{in} * d_{out})$ \\
		\textbf{GGNN} &
		\begin{scriptsize}
		$\begin{aligned}[t]
			& {\boldsymbol{z}}_i^l = \delta ( \textcolor{blue}{\boldsymbol{W}}^z \boldsymbol{s}_i^l + \textcolor{blue}{\boldsymbol{b}}^{sz} + \textcolor{blue}{\boldsymbol{U}}^z \boldsymbol{h}_i^{l} + \textcolor{blue}{\boldsymbol{b}}^{hz}) \\ 
			&\boldsymbol{r}_i^l = \delta ( \textcolor{blue}{\boldsymbol{W}}^r \boldsymbol{s}_i^l+ \textcolor{blue}{\boldsymbol{b}}^{sr} +\textcolor{blue}{\boldsymbol{U}}^r \boldsymbol{h}_i^{l} + \textcolor{blue}{\boldsymbol{b}}^{hr}) \\ 
			&\boldsymbol{h}_i^{l+1} = tanh ( \textcolor{blue}{\boldsymbol{W}} \boldsymbol{s}_i^l + \textcolor{blue}{\boldsymbol{b}}^s + \textcolor{blue}{\boldsymbol{U}} ( \boldsymbol{r}_i^l \odot \boldsymbol{h}_i^{l} + \textcolor{blue}{\boldsymbol{b}}^h))) \\ 
			&\boldsymbol{h}_i^{l+1} = (1 - {\boldsymbol{z}}_i^l) \odot \boldsymbol{h}_i^l +  {\boldsymbol{z}}_i^l \odot \boldsymbol{h}_i^{l+1}
		\end{aligned}$
		\end{scriptsize}
		&
		$O(d_{in} * d_{out})$ \\
		\textbf{GAT}&
		$\boldsymbol{h}_i^{l+1} = \boldsymbol{s}_i^l$ &
		$O(d_{out})$ \\
		\textbf{GaAN} &
		\begin{scriptsize}
			$\begin{aligned}[t]
			&\boldsymbol{g}_i = \textcolor{blue}{\boldsymbol{W}}^l_g  [\boldsymbol{h}_i^{l} \parallel \boldsymbol{s}_{i, 2}^l \parallel \boldsymbol{s}_{i, 3}^l]  \\
			&\boldsymbol{h}_i^{l+1} = \textcolor{blue}{\boldsymbol{W}}^l_o [\boldsymbol{h}_i^l \parallel (\boldsymbol{g}_{i} \odot \boldsymbol{s}_{i, 1}^l) ]
			\end{aligned}$ 
		\end{scriptsize}&
		$O(max(K * d_v + d_{in}, 2 * d_{in} + d_m) * d_{out})$ \\ \bottomrule
	\end{tabular}
\end{footnotesize}
	\caption{Typical graph neural networks and their vertex computation functions. $d_{in}$ and $d_{out}$ are dimensions of the input and output hidden feature vectors, respectively. Blue variables are model parameters to learn. For ChebNet, we report its GNN sub-layer.For Neural FPs, $\textcolor{blue}{\boldsymbol{W}}^{l, |\mathcal{N}(i)|}$ is the weight matrix for nodes with degree
	$|\mathcal{N}(i)|$ at layer $l$.}
	\label{tab:gnn_overview_vertex}
\end{table}

@@

\textbf{GGNN} ()
